A standard for the scholarly citation of archaeological data as an incentive to data sharing

### 2017-12-31 analysis/paper/paper.Rmd
Editor's revisions submitted

To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in Marwick (2017), we include the entire R code used for all the analysis and visualizations contained in this paper in our compendium at http://dx.doi.org/10.17605/OSF.IO/KSRUZ. Also in this version-controlled compendium are the raw data for all the tests reported here. All of the figures and quantitative results presented here can be independently reproduced with the code and data in this repository. In our compendium our code is released under the MIT license, our data as CC-0, and our figures as CC-BY, to enable maximum re-use (for more details, see Marwick 2017).

@@

The _Journal of Archaeological Science_ has had a 'data disclosure' policy since at least 2013 (Editors, 2013) which states that "All data relating to the article must be made available in Supplementary files or deposited in external repositories and linked to within the article". The data availability rate of `r data_are_avail_perc`% reported in our pilot study reflects either weak enforcement of this policy or an incomplete understanding among editors, reviewers, and authors of how to interpret and implement this policy. The prominence of compositional and dating data among the shared data types suggests a least effort strategy by authors sharing data that do not require extensive cleaning and tidying (because the data are generated by instruments rather than entered by hand, such as artefact attribute data). The size of compositional and dating datasets means that they mostly fit easily into tables in the text or supplementary files, and do not require much effort to organize (e.g. very little metadata is required for others to use these data). The small proportion of papers in our sample using or sharing a scripting language suggests that unlike other fields (Eglen et al. 2017), archaeologists do not appear to recognize code for data analysis as a primary research object to share with the research community. Code written by the researcher for even the simplest operations should be shared along with data because code communicates decisions made about the data analysis that can help others understand, analyze, and reuse the data.

@@

```{r citation-time-series}
citation_time_series_plot <-  
events_df_names %>% 
  select(occurred_at) %>% 
  mutate(occurred_at = as.Date(occurred_at)) %>% 
ggplot(aes(occurred_at)) +
  geom_histogram() +
  scale_x_date(date_breaks = "6 month") +
  theme_bw(base_size = 10) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  xlab("Date of citation event") +
  ylab("Number of citations\nper day")

@@

- When preparing research for publication, upload data to a trustworthy repository and cite the data in the manuscript.      
- When writing peer reviews of journal articles, we can tell editors that we require data to be available for papers that we read, and that the manuscripts we review must include citations of datasets analyzed in the manuscript. We can write in our peer reviews that data need to be available for us to make a proper assessment of the manuscript.    
- When reviewing grant applications, we can critically evaluate data management plans, and report on the applicants' past record on data sharing. We can offer helpful suggestions about how to improve data availability.     
- When teaching students, we can teach them how to prepare data for sharing, and how to find and use publicly available datasets to give students experience analyzing and writing about shared data, and to cultivate the expectation that data sharing is a normal part of archaeological research.     
- When evaluating candidates for hiring and promotion, ensure that commitments to data sharing are considered.

@@

Editors
2013.	Guide for Authors. Journal of Archaeological Science. Electronic document, https://web.archive.org/web/20130125222940/https://www.elsevier.com/journals/journal-of-archaeological-science/0305-4403/guide-for-authors, accessed 7 September 7, 2017

@@

@@

>>>>

To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in Marwick (2017), we include the entire R code used for all the analysis and visualizations contained in this paper in our compendium at http://doi.org/10.17605/OSF.IO/KSRUZ. Also in this version-controlled compendium are the raw data for all the tests reported here. All of the figures and quantitative results presented here can be independently reproduced with the code and data in this repository. In our compendium our code is released under the MIT license, our data as CC-0, and our figures as CC-BY, to enable maximum re-use (for more details, see Marwick 2017).

@@

The _Journal of Archaeological Science_ has had a 'data disclosure' policy since at least 2013 (_Journal of Archaeological Science_ 2013) which states that "All data relating to the article must be made available in Supplementary files or deposited in external repositories and linked to within the article". The data availability rate of `r data_are_avail_perc`% reported in our pilot study reflects either weak enforcement of this policy or an incomplete understanding among editors, reviewers, and authors of how to interpret and implement this policy. The prominence of compositional and dating data among the shared data types suggests a least effort strategy by authors sharing data that do not require extensive cleaning and tidying (because the data are generated by instruments rather than entered by hand, such as artefact attribute data). The size of compositional and dating datasets means that they mostly fit easily into tables in the text or supplementary files, and do not require much effort to organize (e.g. very little metadata is required for others to use these data). The small proportion of papers in our sample using or sharing a scripting language suggests that unlike other fields (Eglen et al. 2017), archaeologists do not appear to recognize code for data analysis as a primary research object to share with the research community. Code written by the researcher for even the simplest operations should be shared along with data because code communicates decisions made about the data analysis that can help others understand, analyze, and reuse the data.

@@

```{r citation-time-series}
citation_time_series_plot <-  
events_df_names %>% 
  select(occurred_at) %>% 
  mutate(occurred_at = as.Date(occurred_at)) %>% 
ggplot(aes(occurred_at)) +
  geom_histogram(fill = wes_palette("Rushmore")[4]) +
  scale_x_date(date_breaks = "6 month") +
  theme_bw(base_size = 10) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  xlab("Date of citation event") +
  ylab("Number of citations")

@@

- When preparing research for publication, we can upload data to a trustworthy repository and cite the data in the manuscript.      
- When writing peer reviews of journal articles, we can tell editors that we require data to be available for papers that we read, and that the manuscripts we review must include citations of datasets analyzed in the manuscript. We can write in our peer reviews that data need to be available for us to make a proper assessment of the manuscript.    
- When reviewing grant applications, we can critically evaluate data management plans, and report on the applicants' past record on data sharing. We can offer helpful suggestions about how to improve data availability (e.g. Table \@ref(tab:desirable)).     
- When teaching students, we can instruct them how to prepare data for sharing, how to find and use publicly available datasets to give students experience analyzing and writing about shared data, and to cultivate the expectation among students that data sharing is a normal part of archaeological research.     
- When evaluating candidates for hiring and promotion, we can ensure that commitments to data sharing are valued.

@@

@@

Journal of Archaeological Science
2013	Guide for Authors. Journal of Archaeological Science. Electronic document, https://web.archive.org/web/20130125222940/https://www.elsevier.com/journals/journal-of-archaeological-science/0305-4403/guide-for-authors, accessed 7 September 7, 2017

@@

Tenopir, Carol, Suzie Allard, Kimberly Douglass, Arsev U. Aydinoglu, Lei Wu and Mike Frame
2011   Data Sharing by Scientists: Practices and Perceptions. _PLOS ONE_ 6(6): e21101. DOI:10.1371/journal.pone.0021101, accessed, 7 September 2017
